Depletion capabilities in the OpenMC Monte Carlo particle transport code

### 2020-06-23 manuscript.tex
Add conclusions and acknowledgments

\subsubsection{predictor--Corrector Methods}

@@

Discuss temperature dependence and probability tables in SFR model.

@@

@@

>>>>

\subsubsection{Predictor--Corrector Methods}

@@

A new depletion solver has been implemented in the OpenMC Monte Carlo particle
transport code. The depletion solver is implemented primarily in Python 3 and
interfaces with OpenMC's transport solver through Python bindings to an
underlying C++ API, which enables data transfers to happen in memory rather than
through filesystem I/O. The design of the depletion solver relies on a set of
abstract classes that provide modularity and separation between difference
components, including the time integration methods, the matrix exponential
solver, and the transport operator. In principle, this modularity would allow
the depletion system to be used with another transport code. Rather than relying
on a single time integration method, OpenMC implements several time integration
schemes including the predictor method, standard predictor-corrector methods,
stochastic implicit Euler integrators, and several novel
methods~\citep{josey2017phd, josey2017jcp}. To evaluate the matrix exponential,
OpenMC relies on the incomplete partial fraction form of CRAM by default.

To assess the accuracy of the new depletion solver, comparisons were made
between OpenMC and Serpent on a PWR pincell model and an SFR assembly model.
When the full depletion chain was used in OpenMC, the predictions of
$k_\text{eff}$ agreed to within 20--30 pcm of that calculated by Serpent at all
times in life for both the PWR and SFR model. Actinide concentrations were found
to agree to a fraction of a percent, and most fission product concentrations
were found to agree within 1\%. The few cases where fission product
concentrations differed by a larger amount were found to be related to
differences in the treatment of the energy dependence of FPYs. It is our sense
that there is no consensus in the community as to what is the ``correct'' way of
handling this energy dependence, so it is not possible to opine on whether one
set of results is right or wrong. When the simplified CASL depletion chain was
used in OpenMC, the accuracy of actinide and fission product concentrations did
not change appreciably. The only nuclide for which a significant difference was
observed was $^{85}$Kr; this difference can be explained by a cumulative yield
in the ENDF/B-VII.1 FPY sublibrary that is inconsistent with the independent
yields of precursors~\citep{pigni2015nds}. Because the CASL depletion chain
results in a significantly lower runtime compared to the full depletion chain
while still retaining accuracy, it is recommended for general use by OpenMC
users.

We found that when comparing results on the SFR assembly model, larger
differences in $k_\text{eff}$ and isotopic concentrations were observed when
unresolved resonance region probability tables were turned on. These differences
were found to be related to how the temperature dependence of cross sections is
treated in the unresolved resonance region. In OpenMC, when the temperature of a
material is in between two temperatures at which probability tables are
tabulated, statistical interpolation between the tables is performed. Serpent,
on the other hand, uses probability tables at the lower tabulated temperature.
For the SFR model, this results in lower parasitic neutron absorption in
$^{238}$U, lower production of $^{239}$Pu, and ultimately a lower
$k_\text{eff}$. After depleting for 360 days, there was a 100 pcm difference
between OpenMC and Serpent due to this effect. Although this is not a huge
difference, users who are simulating problems with a non-uniform temperature
distribution should be aware of this behavior.

Finally, we remark that although we did observe some differences between OpenMC
and Serpent, most of the differences were related to deficiencies in the nuclear
data itself. As a whole, the results demonstrate that when the same nuclear data
and same methods are used in two independently developed codes, it is possible
to obtain results in a depletion simulation that match within 10s of pcm on
$k_\text{eff}$ and less than 1\% on isotopic concentrations.

@@

We gratefully acknowledge the OpenMC user community for providing early feedback
and bug reports related to depletion that have helped improve the code. We wish
to acknowledge Jaakko Lepp{\"{a}}nen (VTT) for providing the developmental
version of Serpent 2.1.32 used in this work and for helping to shed light on the
methods used in Serpent. We acknowledge Nicolas Stauff (ANL) for providing
recommendations on a suitable SFR target problem.

@@

We gratefully acknowledge the computing resources provided on Bebop, a
high-performance computing cluster operated by the Laboratory Computing Resource
Center at Argonne National Laboratory. This research made use of the resources
of the High Performance Computing Center at Idaho National Laboratory, which is
supported by the Office of Nuclear Energy of the U.S. Department of Energy and
the Nuclear Science User Facilities under Contract No. DE-AC07-05ID14517.
